---
title: "Gun Control Chi-square"
author: "Robert Saldivar"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This: I'm Robert and I'm updating this markdown file here.

```{r, echo=FALSE}
library(tidyverse)

gc_df <- read_csv("gc_df.csv")
```


WE are asking: Is there a significant difference in recommended gun control action for male and female respondents (should guns be easier, harder, to get?)

This: I'm Nate and I'm updating this markdown file here. 

Robert: data wrangling and counts
```{r}
get_guns <- gc_df%>%
  filter(gender !="NA", make_get_guns !="NA") %>%
  count(gender, make_get_guns) %>%
  spread(make_get_guns, n) %>%
  select(-gender)

rownames(get_guns) <- c("Female","Male")

```

My name is Nate the colaborataor and I am going to be doing chi-square test for independence! 

We ask: "Is there a significant association between sex and opinion on whether guns sould be more difficult, easier, or the same difficulty to get?"

```{r}
# Let us look at the proporitons: 

guns_prop <- prop.table(as.matrix(get_guns), 1 ) #Using the prop.table function with the as.matrix class (, 1 takes all the feamle counts and finds the proportions of respondents; finding proportions by rows )
#Point out actual proportions with Chi-Squared function 

#Run chi-square test for independence (is there a significant association between sex and access to guns) 

guns_x2 <- chisq.test(get_guns)

guns_x2

#H0: There is no difference in the proportions for female and male respondenets 
#Reject the null hypothesis (p < 0.05)!
```
"Recommended government action on gun access differed significantly between male (n = ) and feamle (n =) respondents ($\chi^2$(`rguns_x2$parameter`) = ....., p < 0.0001). Notably, a much higher proportion of female respondents (57.4%) recommend making it more diffucult to get guns, compared to 40.4% for male respondents" 

#guns_x2$stdres #Any values greater than 2 different enough from an expected value would be driving significance 
#What is driving a significant result? 
